---
title: "Homework 7"
author: "Fall 2016, Math 107, Prof. Adam Loy"
date: "Due Tuesday, November 8 by 4 p.m."
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Please complete your homework in this R markdown file. When you are done, please knit the document to Word and print a hard copy for submission. Make sure to change the `author:` so that you get credit for your work!

The below code chunk loads all of the necessary R packages. Note that we will use the `car` package. If you are using your personal computer, you will need to install this package **before** attempting to knit the document.

```{r message=FALSE}
library(ggplot2)
library(mosaic)
library(car)
```


## Problem 1

Energy can be produced from wind using windmills. Choosing a site for a wind farm, the location
of the windmills, can be a multi-million dollar gamble. If wind is inadequate at the site, then the
energy produced over the lifetime of the wind farm can be much less than the cost of building
the operation. Prediction of wind speed at a candidate site can be an important component in the
decision to build or not to build. Since energy produced varies as the square of the wind speed,
even small errors can have serious consequences.

One possible way to predict wind speed at a candidate site is to use wind speed at
a nearby reference site. A reference site is a nearby location where the wind speed should be
similar to the candidate site. Airports sometimes serve as reference sites, but in our case, 
the reference data come from the National Center for Environmental Modeling (NCAR, 2013). Both sites
are in the northern part of South Dakota.

The data set `windmill.csv` contains measurements of wind speed (in meters per second, m/s) 
at the candidate site (CSpd) and at a accompanying reference site (RSpd). 

First, run the below code chunk to load the `windmill.csv` data set.

```{r}
windmill <- read.csv("data/windmill.csv")
```

#### Part a.

Identify the explanatory and response variables for this problem.

*Write your answer here, replacing this text.*

#### Part b.

Create a scatterplot of the wind speed at the candidate site vs. the wind speed at the accompanying reference site.

```{r fig.height=3, fig.width=5}
# INSERT YOUR CODE HERE
```


#### Part c.

Describe the relationship between the wind speed at the candidate site and the wind speed at the accompanying reference site.

*Write your answer here, replacing this text.*


#### Part d.

Calculate the correlation between the wind speed at the candidate site and the wind speed at the accompanying reference site.

```{r}
# INSERT YOUR CODE HERE
```

#### Part e.

Fit the least squares regression line and print a summary of the model.

```{r}
# INSERT YOUR CODE HERE
```

#### Part f.

Report the fitted least squares regression equation.

*Write your answer here, replacing this text.*

#### Part g.

Interpret the estimate of the slope within the context of the problem.

*Write your answer here, replacing this text.*


#### Part h.

Interpret the estimate of the y-intercept within the context of the problem.

*Write your answer here, replacing this text.*


#### Part i.

Calculate a 95% bootstrap confidence interval for the slope. 

```{r}
# INSERT YOUR CODE HERE
```


#### Part j. 

Interpret the interval you just created in the context of the problem.

*Write your answer here, replacing this text.*


#### Part k.

Based on this interval, is there statistically significant evidence that the slope is not zero? Explain.

*Write your answer here, replacing this text.*


#### Part l.

What proportion of the variability in wind speed at the candidate site is explained by the regression model?

*Write your answer here, replacing this text.*


#### Part m.

Predict the wind speed at the candidate site on a day when the wind speed at the reference site is 10 m/s.

*Write your answer here, replacing this text.*


#### Part n. 

Would you be comfortable using your model to predict the wind speed at the candidate site given the wind speed at the reference site is 30m/s? Explain.

*Write your answer here, replacing this text.*


#### Part o.

Create a residual plot for your model. Is there any apparent pattern in the residual plot? What does this indicate about the linearity and constant variability?


```{r fig.height=3, fig.width=5}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


#### Part p.

Create a histogram of the residuals. Describe the distribution of the residuals and what this indicates about the appropriateness of the model?


```{r fig.height=3, fig.width=5}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*
