---
title: "Homework 2"
author: "Fall 2016, Math 107, Prof. Adam Loy"
date: "Due Tuesday, September 20 by 4:00 p.m."
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1
The bar chart and the pie chart below show the distribution of pre-existing medical 
conditions of children involved in a study on the optimal duration of 
antibiotic use in treatment of tracheitis, which is an upper respiratory 
infection.

```{r eval=FALSE, include=FALSE}
<<include=FALSE>>=
# load openintro package for colors ---------------------------------
library(openintro)

# create data -------------------------------------------------------
library(forcats)
conditions = c(rep("Prematurity", 33),
               rep("Neuromuscular", 10),
               rep("Cardiovascular", 16),
               rep("Genetic/metabolic", 6),
               rep("Respiratory", 13),
               rep("Trauma", 10),
               rep("Gastrointestinal", 2),
               rep("Immunocompromised", 2)
               )

conditions <- fct_relevel(conditions, "Prematurity", "Cardiovascular", "Respiratory", "Trauma", "Neuromuscular", "Genetic/metabolic", "Immunocompromised", "Gastrointestinal")

conditions <- fct_rev(conditions)

conditions_df <- data.frame(conditions = conditions)
rm(conditions)



# bar chart -----------------------------------------------------------
library(ggplot2)
library(viridis) # nicer color palettes
bar <- 
  ggplot(data = conditions_df, mapping = aes(x = conditions, fill = conditions)) + 
  geom_bar() + 
  coord_flip() +
  scale_fill_viridis(discrete=TRUE)

# pie chart ---------------------------------------------------------
pie <- 
  ggplot(data = conditions_df, mapping = aes(x = factor(1), fill = conditions)) + 
  geom_bar(width=1) + 
  coord_polar(theta = "y") +
  scale_fill_viridis(discrete=TRUE) +
  labs(x = NULL, y = NULL)

require(gridExtra)
grid.arrange(bar, pie, ncol = 2)
```

a. What features are apparent in the bar chart but not in the pie chart?
b. What features are apparent in the pie chart but not in the bar chart?
c. Which graph would you prefer to use for displaying these categorical data?


## Problem 2

In this problem we will explore a data set containing the velocity (in km/sec) of 82 planets.

```{r echo=FALSE, fig.height=3, fig.width=7, warning=FALSE}
library(ggplot2)
library(MASS)
library(gridExtra)
galaxy <- data.frame(velocity = galaxies)
p1 <- ggplot(data = galaxy, mapping = aes(velocity)) +
  geom_histogram(binwidth = 1000, color = "gray20") + 
  labs(x= "velocity (km/sec)") +
  theme(axis.text.x = element_text(size = rel(.8))) +
  theme_bw()

p2 <- ggplot(data = galaxy, mapping = aes(velocity)) +
  geom_density() + 
  theme(axis.text.x = element_text(size = rel(.8)))

p3 <- ggplot(data = galaxy, mapping = aes("var", velocity)) +
  geom_boxplot() +
  labs(x = "", y = "velocity (km/sec)") +
  scale_x_discrete(breaks = NULL) +
  theme_bw()

grid.arrange(p1, p3, ncol = 2)
```

a. Describe the shape of the distribution of velocities. Specifically address the number of modes, symmetry/skew, and any unusual features present.
b. What features are apparent in the histogram but not in the boxplot?
c. Which graph would you prefer to use for displaying these data? Justify your answer.

## Problem 3

In this problem we will explore the length of movies released

```{r echo=FALSE, fig.height=3, fig.width=5, warning=FALSE}
data("movies", package="bigvis")

ggplot(data = movies, mapping = aes(x = length)) + 
  geom_histogram(binwidth = 1) +
  labs(x = "Movie lengths (minutes)", y = "Count") +
  xlim(0, 180) +
  theme_bw()
```

a. Describe the shape of the distribution of velocities. Specifically address the number of modes, symmetry/skew, and any unusual features present.
b. Below is a density plot displaying the same data. What features are easier to see in the density plot? What features are easier to see in the histogram?

```{r echo=FALSE, fig.height=3, fig.width=5, warning=FALSE}
ggplot(data = movies, mapping = aes(x = length)) + 
  geom_density() +
  labs(x = "Movie lengths (minutes)", y = "Density") +
  xlim(0, 180) +
  theme_bw()
```

c. Below is a side-by-side (i.e. faceted) histogram of movie lengths based on whether the film is classified as a "short" film. Have films been classified as a "short" in a consistent way? Explain briefly.

```{r echo=FALSE, fig.height=4.5, fig.width=5, warning=FALSE, message=FALSE}
library(dplyr)
movies <- movies %>%
  mutate(short_lab = ifelse(Short==0, "Feature", "Short"))

ggplot(data = movies, mapping = aes(x = length)) + 
  geom_histogram(binwidth = 1) +
  facet_wrap(~short_lab, ncol=1) +
  labs(x = "Movie lengths (minutes)", y = "Count") +
  xlim(0, 180) +
  theme_bw()
```


b. The below plot displays the same data, but uses overlaid density plots instead of histograms. What features are easier to see with the density plots? What features are easier to see in the histograms?

```{r echo=FALSE, fig.height=3, fig.width=5, warning=FALSE, message=FALSE}
library(viridis) # nicer color palettes
ggplot(data = movies, mapping = aes(x = length, fill = short_lab)) + 
  geom_density(alpha = 0.5) +
  labs(x = "Movie lengths (minutes)", y = "Density") +
  xlim(0, 180) +
  scale_fill_viridis('Type', discrete = TRUE) +
  theme_bw()
```


## Problem 4

```{r echo=FALSE, fig.width=10, fig.height=10, cache=TRUE, warning=FALSE}
library(extracat)
library(GGally)

data("olives")
ggpairs(olives[,3:10], 
        lower = list(continuous = wrap("points", alpha = 0.3)),
                     title = "Olive Oil Fatty Acids") +
  theme_bw()
```

```{r echo=FALSE, fig.width=10, fig.height=10, cache=TRUE, warning=FALSE}
library(ggthemes)
p <- ggpairs(olives[,2:10], 
        mapping = ggplot2::aes(color = Region),
        lower = list(continuous = wrap("points", alpha = 0.3)),
                     title = "Olive Oil Fatty Acids") +
  theme_bw()
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_colorblind() +
        scale_color_colorblind()  
  }
}

p
```



