---
title: "Homework 2"
author: "Fall 2016, Math 107, Prof. Adam Loy"
date: "Due Tuesday, September 20 by 4:00 p.m."
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1
The bar chart and the pie chart below show the distribution of pre-existing medical 
conditions of children involved in a study on the optimal duration of 
antibiotic use in treatment of tracheitis, which is an upper respiratory 
infection.

```{r eval=FALSE, include=FALSE}
<<include=FALSE>>=
# load openintro package for colors ---------------------------------
library(openintro)

# create data -------------------------------------------------------
library(forcats)
conditions = c(rep("Prematurity", 33),
               rep("Neuromuscular", 10),
               rep("Cardiovascular", 16),
               rep("Genetic/metabolic", 6),
               rep("Respiratory", 13),
               rep("Trauma", 10),
               rep("Gastrointestinal", 2),
               rep("Immunocompromised", 2)
               )

conditions <- fct_relevel(conditions, "Prematurity", "Cardiovascular", "Respiratory", "Trauma", "Neuromuscular", "Genetic/metabolic", "Immunocompromised", "Gastrointestinal")

conditions <- fct_rev(conditions)

conditions_df <- data.frame(conditions = conditions)
rm(conditions)



# bar chart -----------------------------------------------------------
library(ggplot2)
library(viridis) # nicer color palettes
bar <- 
  ggplot(data = conditions_df, mapping = aes(x = conditions, fill = conditions)) + 
  geom_bar() + 
  coord_flip() +
  scale_fill_viridis(discrete=TRUE)

# pie chart ---------------------------------------------------------
pie <- 
  ggplot(data = conditions_df, mapping = aes(x = factor(1), fill = conditions)) + 
  geom_bar(width=1) + 
  coord_polar(theta = "y") +
  scale_fill_viridis(discrete=TRUE) +
  labs(x = NULL, y = NULL)

require(gridExtra)
grid.arrange(bar, pie, ncol = 2)
```

a. What features are apparent in the bar chart but not in the pie chart?
b. What features are apparent in the pie chart but not in the bar chart?
c. Which graph would you prefer to use for displaying these categorical data?


## Problem 2

HISTOGRAMS/DENSITY PLOTS

## Problem 3

SIDE-by-SIDE plots

## Problem 4

SCATTERPLOT

