---
title: 'Lab 3: Introduction to Hypothesis Testing'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## 0. Intro

Welcome to Lab 4!

Today's lab will continue to explore the idea of hypothesis testing permutation tests, but this lab will focus on the situation where we have two categorical variables.

### 0.1 Administrative details
Lab submissions are due by Friday, October 14 by 4:00 p.m. To submit your assignment, please upload both the .Rmd and .html files on Moodle.

### 0.2 Setup
You can download the .Rmd file for this lab and the data from the course [webpage](http://math107-lu.github.io/) as a zip file. If you are using a Mac, then you will need to use a browser other than Safari for the download. 

If you are using the RStudio server, then you can upload the entire zip folder directly onto the server.

We will use the following packages during this lab. Make sure that you have downloaded all of them before running the commands.

```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(CarletonStats)
```


## 1. Comparing Two Proportions

## 1.1. The data

Is yawning contagious? Conventional wisdom says yes: When we see someone else yawn, we're prone to let out a yawn ourselves, but will data support this claim if we put it to a scientific test?

The folks at *MythBusters*, a popular television program on teh Discovery Channel, investigated this issue by using a hidden camera. Fifty people attending a local flea market were recruisted to participate. Subjects were ushered, one at a time, into one of three rooms by co-host Kari. She yawned (planting a yawn "seed") as she ushered subjects into two of the rooms, and for the other room she did not yawn. The researchers decided in advance, with a random mechanism, which subjects went to which room. As time passed, the researchers watched to see which subjects yawned.

The data for this study are found in the `yawning.csv` data file.


**Question 1.** Is this an observational study or an experiment? Briefly justify your answer.

**Question 2.** Why did the researchers randomly assign subjects to the "yawn seed" group?

Now that we understand the study design, we can load the data set.

```{r}
yawning <- read.csv("data/yawning.csv")
summary(yawning)
```

The rows describe the individual subjects. The columns describe:

- **YawnSeed**: Whether or not the "yawn seed" was planted (`Control` or `Seeded`)
- **Response**: Whether or not the subject yawned (`NoYawn` or `Yawn`)

The researchers wish to determine whether or not yawning is contagious.

**Question 3.** State the null and alternative hypotheses using words and notation.


### 1.2. Exploring the association

Before conducting a statistical test for whether or not yawning is contagious, we will explore the possible association.

**Question 2.** Create a bar chart of whether or not the subject yawned (`Response`) segmented by whether or not the "yawn seed" was planted (`YawnSeed`).

**Question 4.** Fill in the following table:

                     | Yawn seed planted | Yawn seed not planted | Total
                     |-------------------|-----------------------|----------------
Subject yawned       |-------------------|-----------------------|----------------
Subject did not yawn |-------------------|-----------------------|----------------
Total                |-------------------|-----------------------|----------------


**Question 5.** Calculate the proportion of subjects in the "yawn seed" group that yawned.

**Question 5.** Calculate the proportion of subjects in the control group that yawned.

**Question** Describe the apparent association between the variables revealed by the plot and summary statistics.

## 2. Using Random Permutations



## References

[John Rauser's keynote address at Strata + Hadoop 2014](https://youtu.be/5Dnw46eC-0o) serves as great motivation to learn about permutation tests, and reference bootstrapping. (We will do that too!)

Lefèvre, Thierry, et al. "Beer consumption increases human attractiveness to malaria mosquitoes." *PloS one* 5.3 (2010): e9546. http://dx.doi.org/10.1371/journal.pone.0009546