---
title: "Creating Graphs with ggplot2"
author: "Lab 1, Fall 2016"
output:
  html_document:
    fig_height: 2
    fig_width: 5
  word_document:
    fig_height: 2
    fig_width: 5
---

### 1. Introduction
It is often necessary to create graphics to effectively communicate key patterns within a data set. While many software packages allow the user to make basic plots, it can be challenging to create plots that are customized to address a specific idea. While there are numerous ways to create graphs, this tutorial will focus on the R package `ggplot2`, created by Hadley Wickham.


In this lab, we will focus on building plots in layers using the `ggplot2` package. This approach uses a particular grammar inspired by Leland Wilkinson's landmark book, *The Grammar of Graphics*, that focused on thinking about, reasoning with and communicating with graphics. It enables layering of independent components to create custom graphics for tidy data sets.

To begin, load the `ggplot2` package using the following command
```{r message = FALSE}
library(ggplot2)
```

If you have not yet installed the package, you can do so with the following command

```{r eval=FALSE}
install.packages("ggplot2")
```


**Data**: In this tutorial, we will use the  [AmesHousing](http://www.amstat.org/publications/jse/v19n3/decock.pdf) data set, which provides information on the sales of individual residential properties in Ames, Iowa from 2006 to 2010. The data set contains 2930 observations, and a large number of explanatory variables involved in assessing home values. A full description of this data set can be found [here](http://www.amstat.org/publications/jse/v19n3/Decock/DataDocumentation.txt).

If you have saved the data file in the data folder of your R Studio project, then you can load it using the following command. Alternatively, you can use "Import Dataset" button in the Environment tab.

```{r}
AmesHousing <- read.csv("data/AmesHousing.csv")
# str(AmesHousing)
```

### 2. Building your first plot

To see how plots are constructed, let's first create a histogram of the sales price (`SalePrice`) of homes in Ames, IA. To do this, we will use the `ggplot` function which expects at a bare minimal as arguments:

* the data frame where the variables exist (the `data` argument) and
* the names of the variables to be plotted (the `mapping` argument).

The names of the variables will be entered into the aes function as arguments where aes stands for "aesthetics".

With that in mind, we start by creating the base layer, or the backdrop on which we will layer the elements of our histogram.

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice))
```

We next proceed by adding a layer---hence, the use of the + symbol---to the plot to produce a histogram.

Note: You are encouraged to enter **Return** on your keyboard after entering the `+`. As we add more and more elements, it will be nice to keep them indented as you see below. Note that this will not work if you begin the line with the `+`.

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice)) +
  geom_histogram()
```

Here, we are warned that the histogram was been drawn with 30 bins, a rather arbitrary choice. We can customize our histogram by adding a `binwidth` argument to the `geom_histogram` command:

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice)) +
  geom_histogram(binwidth=10000)
```

We can also add some color to the plot by invoking the fill and color arguments.

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice)) +
  geom_histogram(binwidth=10000, fill = "steelblue2", color = "black")
```

Currently, the x-axis label does not contain units. To change the axis labels we add a layer to our plot

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice)) +
  geom_histogram(binwidth=10000, fill = "steelblue2", color = "black") +
  labs(x = "Sale Price (in dollars)")
```

Note that we can also tweak the y-axis label with the addition of a `y` argument in the `labs` function.

Finally, we can easily add a title to our plot with the addition of another layer

```{r}
ggplot(data = AmesHousing, mapping = aes(x = SalePrice)) +
  geom_histogram(binwidth=10000, fill = "steelblue2", color = "black") +
  labs(x = "Sale Price (in dollars)") +
  ggtitle("Sales Prices of Homes in Ames, IA")
```


Throughout the remainder of this lab we will explore how to create and customize the basic plot types:

* histograms,
* bar charts,
* density plots,
* scatterplots,
*  and side-by-side plots.