---
title: "Homework 9"
author: "Fall 2016, Math 107, Prof. Adam Loy"
date: "Due Wednesday, November 16 at the beginning of class"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Please complete your homework in this R markdown file. When you are done, please knit the document to Word and print a hard copy for submission. Make sure to change the `author:` so that you get credit for your work!

The below code chunk loads all of the necessary R packages.

```{r message=FALSE}
library(ggplot2)
library(mosaic)
```

## Problem 1
An independent random sample is selected from an approximately normal population with unknown standard deviation. Find the degrees of freedom and use R to find the critical t-value ($t^*$) for the given sample size and confidence level.

#### Part a.
 $n=6$,$\rm{CL}=90\%$ 
 
```{r}
# INSERT YOUR CODE HERE
```

#### Part b.
$n=21$,$\rm{CL}=98\%$

```{r}
# INSERT YOUR CODE HERE
```

#### Part c.
$n=29$,$\rm{CL}=95\%$ 

```{r}
# INSERT YOUR CODE HERE
```

#### Part d.
$n=12$,$\rm{CL}=99\%$

```{r}
# INSERT YOUR CODE HERE
```


## Problem 2
An independent random sample is selected from an approximately normal population with an unknown 
standard deviation. Find the p-value for the given set of hypotheses and $T$ 
test statistic. Also determine if the null hypothesis would be rejected 
at $\alpha = 0.05$.

#### Part a.
$H_A: \mu > \mu_0$, $n = 11$, $T = 1.91$

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


#### Part b.
$H_A: \mu < \mu_0$, $n = 17$, $T = -3.45$

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


#### Part c.
$H_A: \mu \ne \mu_0$, $n = 7$, $T = 0.83$

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


#### Part d.
$H_A: \mu > \mu_0$, $n = 28$, $T = 2.13$

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


## Problem 3
Researchers interested in lead exposure due to car exhaust sampled the blood 
of 52 police officers subjected to constant inhalation of automobile exhaust 
fumes while working traffic enforcement in a primarily urban environment. The 
blood samples of these officers had an average lead concentration of 
109.4 $\mu$g/l  and a SD of 36.4 $\mu$g/l; a previous study of individuals 
from a nearby suburb, with no history of exposure, found an average blood 
level concentration of 35 $\mu$g/l.[^1]

The data set `leadexposure.csv` contains the measured lead concentration found in the blood samples. 

First, run the below code chunk to load the `leadexposure.csv` data set.

```{r}
lead <- read.csv("data/leadexposure.csv")
```


#### Part a.
Write down the hypotheses that would be appropriate for testing if the 
police officers appear to have been exposed to a higher concentration of lead.

*Write your answer here, replacing this text.*


#### Part b.
Explicitly state and check the conditions/assumptions necessary for inference on these 
data.

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


#### Part c.
Test the hypothesis that the downtown police officers have a higher lead 
exposure than the group in the previous study. Interpret your results in context.

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*

#### Part d.
Based on your preceding result, without performing a calculation, would 
a 99% confidence interval for the average blood concentration level of police 
officers contain 35 $\mu$g/l?

*Write your answer here, replacing this text.*


#### Part e.
Verify your answer to part (d) by constructing a 99% confidence interval 
for the average blood concentration level of police officers. Provide an interpretation
of this interval in the context of the problem.

```{r}
# INSERT YOUR CODE HERE
```

*Write your answer here, replacing this text.*


[^1]: WI Mortada et al. "Study of lead exposure from automobile exhaust as a risk for nephrotoxicity among traffic policemen." In: *American journal of nephrology* 21.4 (2000), pp. 274–279.